---
title: "t test introduction"
format: html
editor: visual
---

## Let's explore the 'one-sample t test'


```{r}
# remotes::install_github("EvaMaeRey/ggt.test")
library(ggt.test)
library(tidyverse)
# package repo:  github.com/EvaMaeRey/ggt.test

```

# Scenario 3: Hours of sleep.

Students are asked to report their hours of sleep from the night previous.  Is there evidence that on average, there sleep differs from the recommended 8 hours? 

How many students are in the sample?

> 22

What is the mean number of hours? (balancing point)

>

What is the null hypothesis? (put it on the plot) 

> 8

Does it look like our campus population on average get 8 hours of sleep, or is what we observe inconsistn with this proposed mean, $\mu = 8$ 

> looks inconsistent with the proposed 8 hrs average.

```{r}
data_hrs_sleep |> 
  pull(hrs_sleep) |> 
  mean()


data_hrs_sleep |>
  ggplot() + 
  aes(x = hrs_sleep) + 
  geom_mean() + 
  geom_rug() + 
  geom_histogram() + 
  geom_support() + 
  stamp_mean(8) + 
  geom_mean_label() + 
  stamp_mean_label(8) + 
  geom_tdist_null(8, tails = "both")


data_hrs_sleep |>
  ggplot() + 
  aes(x = hrs_sleep) + 
  geom_mean() + 
  geom_rug() + 
  geom_histogram() + 
  geom_support() + 
  stamp_mean(7.25) + 
  geom_mean_label() + 
  stamp_mean_label(7.25) + 
  geom_tdist_null(7.25, tails = "both")
```


# Scenario 2: Do 'normal students' have a mean of grit scores that are consistent west points mean of 3.75

What is the 'n'

>

What is the sample mean?

>

What is the asserted mean (null) $\mu = ?$

>

Does the sample mean (DU students), look like the asserted mean (West Point students), i.e. can I reject the null hypothesis?

>

```{r}
grit_undergrad_data



```


