---
output: github_document
---
   
<!-- README.md is generated from README.Rmd. Please edit that file --> 

```{r, include = FALSE}
knitr::opts_chunk$set(      
  collapse = TRUE,
  warning = FALSE,  
  comment = "#>",
  message = F
)        

```

## {ggprop.test} is a micro package to teach the logic of the prop test.


```{r}
library(tidyverse)

chickwts |> head()
```

```{r, eval = T, echo = T}
library(ggprop.test)
snapshot <- ggplyr::intercept

chickwts |>
  ggplot() + 
  aes(x = weight) +
  geom_rug() +
  geom_histogram() + snapshot("p1") +
  geom_support() + snapshot("p2") + NULL
  # geom_mean() + 
  # geom_mean_label() + snapshot("p3") +
  # stamp_mean(x = ) + 
  # stamp_mean_label(x = ) + snapshot("p4")# +
  # stamp_eq_norm_prop() +
  # geom_normal_prop_null() + snapshot("p5") +
  # geom_normal_prop_null_sds() + snapshot("p6")
```

If we discuss each of the snapshot points, we could write something like this:

```{r, eval = F, fig.width=12, fig.height=12}
library(patchwork)
(p1 + p2) / 
(p3 + p4) /
(p5 + p6)  + 
  patchwork::plot_annotation(
    tag_levels = 1,
    title = "A prop test 'graphical poem' addressing the question: 
    
        For survey of 161 individuals on willingness to serve as 
        organ doners in the case of an accident, is there evidence
        that responses rate differs from a 50/50 split, 
        when 53 individuals respond 'no' and 108 individuals respond 'yes'?
      ",
    subtitle = "1. Draw bar chart (stacks) for each category.
2. Draw theoretical limits for proportion (0 to 1).
3. Calculate proportion and place (also the balancing point for stacks!).
4. Place null (look at the initial question, 'differs from 50/50' in this case)
5. Draw distribution for NULL, where SD = sqrt((p * (1-p))/n)
6. Calc z-score - where does observed prop fit into NULL distributions 
   (i.e. how many standard deviations fit between .5 and .67)"
                             ) &
  ggchalkboard:::theme_blackboard(base_size = 14) 


```

---

# What's inside and how did we get here?
   
## step 00. cloning statexpress functions

Some convenience functions from {statexpress} are used, because we want this to be a bit more self-contained at this point, so we just clone them for now.  statexpress is evolving and is not on CRAN.

<details statexpress functions>

```{r statexpress}
qlayer <- function (mapping = NULL, data = NULL, geom = ggplot2::GeomPoint, stat = StatIdentity, 
    position = position_identity(), ..., na.rm = FALSE, show.legend = NA, 
    inherit.aes = TRUE) 
{
    ggplot2::layer(data = data, mapping = mapping, geom = geom, 
        stat = stat, position = position, show.legend = show.legend, 
        inherit.aes = inherit.aes, params = rlang::list2(na.rm = na.rm, 
            ...))
}

qstat <- function (compute_group = ggplot2::Stat$compute_group, ...) 
{
    ggplot2::ggproto(NULL, Stat, compute_group = compute_group, 
        ...)
}

qstat_panel <- function (compute_panel, ...) 
{
    ggplot2::ggproto(NULL, Stat, compute_panel = compute_panel, 
        ...)
}


proto_update <- function (`_class`, `_inherit`, default_aes_update = NULL, ...) 
{
    if (!is.null(default_aes_update)) {
        default_aes <- aes(!!!modifyList(`_inherit`$default_aes, 
            default_aes_update))
    }
    ggplot2::ggproto(`_class` = `_class`, `_inherit` = `_inherit`, 
        default_aes = default_aes, ...)
}

qproto_update <- function (`_inherit`, default_aes_update = NULL, ...) 
{
    proto_update(NULL, `_inherit`, default_aes_update = default_aes_update, 
        ...)
}

```

</details>



Now let's see the compute...

```{r compute_for_prop_story, echo = T, message=F, warning=F}
library(tidyverse)

# 3. layer add x span
compute_balance <- function(data, scales){
  
  data %>% 
    dplyr::summarise(min_x = min(x, na.rm = T),
              xend = max(x, na.rm = T),
              y = 0,
              yend = 0) %>% 
    dplyr::rename(x = min_x)
  
}


# 4. layer add balancing point 
compute_xmean_at_y0 <- function(data, scales){
  
  data %>% 
    dplyr::summarise(x = mean(x),
              y = 0, 
              label = "^") 
  
}

# 5. layer add balancing point value label
compute_xmean_at_y0_label <- function(data, scales){
  
  data %>% 
    dplyr::summarise(x = mean(x),
              y = 0, 
              label = after_stat(round(x, 2))) 
  
}



# 6. Add 'point' for asserted balancing point (null)
compute_panel_prop_asserted <- function(data, scales, null){
  
  # stamp type layer - so ignore input data
  data.frame(y = 0, 
             x = null,
             label = "^"
             )
  
}

# 6. Add label for asserted balancing point (null)
compute_panel_prop_asserted_label <- function(data, scales, null){
  
  # stamp type layer - so ignore input data
  data.frame(y = 0, 
             x = null,
             label = round(null, 2)
             )
  
}




# 7. normal distribution based on null and n
compute_dist_t <- function (data, scales, height = NULL) 
{
  
  height = height %||% 1.5*nrow(data)/30
  
    seq(-5, 5, 0.01) %>% 
    tibble(x = .) %>% 
    mutate(y_density = dt(x, df = length(data$x) - 
        1)) %>%
    mutate(y = height*y_density/max(y_density)) %>% 
    mutate(x = x * (sd(data$x)/sqrt(length(data$x))) + 
        mean(data$x))
}

# chickwts |> 
#   rename(x = weight) |>
#   compute_dist_t(null = 300) |> head()



# 8. normal distribution mean and sds based on null and n
compute_dnorm_prop_sds <- function(data, scales, null,
  dist_sds = -4:4){
  
  n <- data |> dplyr::count(.by = x) |> dplyr::pull(n) |> sum()
  
  sd = sqrt(null * (1 - null)/n) # sd of the null distribution
  
  q <- dist_sds * sd + null
  
  data.frame(x = q) %>%
    dplyr::mutate(height = dnorm(q, sd = sd, mean = null)) %>%
    dplyr::mutate(height_max = dnorm(0, sd = sd, mean = 0)) %>%
    dplyr::mutate(y = .55*n*height/height_max) %>% # This is a bit fragile...
    dplyr::mutate(xend = x,
           yend = 0)

}  
```

```{r stamp_eq_norm_prop}
#' @export
stamp_eq_norm_prop <- function(x = I(.125),
    y = I(.8), size = 3.5){
  
  annotate(
    "text",
    x = x,
    y = y,
    label = latex2exp::TeX("s = \\sqrt{\\frac{p*(1-p)}{n}}", output = "character"),
    parse = TRUE,
    size = size
  )

}


```






# Step 1. Use the compute in a rendered plot

Sketch mode - where you don't actually define functions, we just have proposals commented out.


```{r ttest_poem, echo = T, message = F, warning=F}
chickwts |>
  ggplot() +
  aes(x = weight) +
  geom_rug() + 
  geom_histogram() +
  # 3. geom_xrange, show scale, range at y is zero
  qlayer(geom = GeomSegment,
         stat = qstat_panel(compute_balance)) +
  # 4. geom_mean, show prop, i.e. balancing point
  qlayer(geom = qproto_update(GeomText, aes(size = 6, vjust = 1)),
         stat = qstat_panel(compute_xmean_at_y0)) +
  # scale_x_discrete(palette = scales::pal_manual(0:1)) +
  # 5. geom_mean_label, labeling prop, balancing point
  qlayer(geom = qproto_update(GeomLabel, aes(fill = from_theme(colour %||% paper), label.size = NA, vjust = 0)),
         stat = qstat_panel(compute_xmean_at_y0_label))  +
  # 6. stamp_mean, assertion, point
  qlayer(geom = qproto_update(GeomText, aes(size = 6, vjust = 1, color = from_theme(colour %||% accent))),
         stat = qstat_panel(compute_panel_prop_asserted),
         null = 300) +
  # 7. stamp_mean_label, assertion, label
  qlayer(geom = qproto_update(GeomLabel,
                              aes(fill = from_theme(colour %||% paper),
                                  label.size = NA, vjust = 0,
                                  color = from_theme(colour %||% accent))),
         stat = qstat_panel(compute_panel_prop_asserted_label),
         null = 300) +
  stamp_eq_norm_prop() +
  # 8. geom_norm on prop plot
  qlayer(geom = qproto_update(GeomArea, aes(alpha = .2)),
         stat = qstat(compute_dist_t),
         null = 300, fill = "blue") +
  ggxmean:::geom_tdist() +
  # 9. geom_prop_norm w/ sd marks
   qlayer(geom = qproto_update(GeomSegment, aes(linetype = "dotted")),
          stat = qstat_panel(compute_dnorm_prop_sds)) +
  labs(title = "Is there statistical evidence that chick weight has dropped\nfrom it's historic average of 300?") 

```

# And then define the functions...

<details>

```{r prop_poem_functions, echo = T, message = F, warning=F}


# #' @export
# geom_stack <- function(...){
#   qlayer(geom = qproto_update(ggplot2::GeomTile, ggplot2::aes(color = "white")), 
#          stat = qstat(compute_group_bricks), 
#          ...)
#   } 

# #' @export
# geom_stack_label <- function(...){
#   qlayer(geom = qproto_update(ggplot2::GeomText, ggplot2::aes(vjust = 0)), 
#          stat = qstat(compute_group_count), 
#          ...)
#   } 

#' @export
geom_support <- function(...){
  qlayer(geom = ggplot2::GeomSegment, 
         stat = qstat_panel(compute_balance), 
         ...)
  }


scale_x_prop <- function(...){ 

  scale_x_discrete(palette = scales::pal_manual(0:1), ...) 
  
}

#' @export
geom_prop <- function(...){
  list(
  qlayer(geom = qproto_update(ggplot2::GeomText, ggplot2::aes(size = 6, vjust = 1)),
         stat = qstat_panel(compute_xmean_at_y0),
         ...),
  scale_x_prop()
  )
  }

#' @export
geom_prop_label <- function(...){ 
  qlayer(geom = qproto_update(ggplot2::GeomLabel, 
                              ggplot2::aes(fill = ggplot2::from_theme(colour %||% paper), label.size = NA, vjust = 0)),
         stat = qstat_panel(compute_xmean_at_y0_label), 
         ...) 
  }

#' @export
stamp_prop <- function(...){ 
  qlayer(geom = qproto_update(ggplot2::GeomText, 
                              ggplot2::aes(size = 6, 
                                           vjust = 1, 
                                           color = ggplot2::from_theme(colour %||% accent))),
         stat = qstat_panel(compute_panel_prop_asserted), 
         ...)
  }
  
#' @export  
stamp_prop_label <- function(...){  
  qlayer(geom = qproto_update(ggplot2::GeomLabel, 
                              ggplot2::aes(fill = ggplot2::from_theme(colour %||% paper), 
                                  label.size = NA, vjust = 0, 
                                  color = ggplot2::from_theme(colour %||% accent))),
         stat = qstat_panel(compute_panel_prop_asserted_label), 
         ...)
  }

#' @export
geom_normal_prop_null <- function(...){
  qlayer(geom = qproto_update(ggplot2::GeomArea, ggplot2::aes(alpha = .2)),
         stat = qstat_panel(compute_dist_t), 
         ...)
  } 

#' @export
geom_normal_prop_null_sds <- function(...){
   qlayer(geom = qproto_update(ggplot2::GeomSegment, ggplot2::aes(linetype = "dotted")),
          stat = qstat_panel(compute_dnorm_prop_sds), 
          ...)
  }




```

</details>


# Done!  See below for all layers, and the introduction of packag for potential break-points along the prop test train of thought.

```{r, eval = F}
chickwts |>
  ggplot() + 
  aes(x = weight) +
  geom_rug() + 
  geom_histogram() + 
  geom_support() + 
  geom_prop() + 
  geom_prop_label() + 
  stamp_prop() + 
  stamp_prop_label() + 
  geom_normal_prop_null() +
  geom_normal_prop_null_sds() + 
  stamp_eq_norm_prop()
```


```{r, eval = F}
# ?prop.test
prop.test(x = sum(donor$Choice == "donor"), # n true
          n = length(donor$Choice), # num observations
          p = .5) # null
```


---

# Minimal Packaging


```{r, eval = F}
knitrExtra::chunk_to_dir("statexpress")
knitrExtra::chunk_to_dir("compute_for_prop_story")
knitrExtra::chunk_to_dir("prop_poem_functions")
knitrExtra::chunk_to_dir("stamp_eq_norm_prop")
```

```{r, eval = F}
usethis::use_package("dplyr")
usethis::use_package("ggplot2")

devtools::document()
devtools::check(".")
devtools::install(pkg = ".", upgrade = "never")
```


<!-- # epilogue... -->


<!-- Another approach that is less concerned with being a bridge to ggplot2 layer extension...?   -->

<!-- ```{r, eval = F, echo = F} -->
<!-- geom_support <- function(...){geom_segment(data = compute_balance, ...)} -->
<!-- geom_prop <- function(...){geom_point(data = compute_xmean_at_y0, label = "^", ...)} -->

<!-- donor |> -->
<!--   ggplot() +  -->
<!--   aes(x = decision) + -->
<!--   geom_support() + -->
<!--   geom_prop() -->
<!-- ``` -->
